---
title: 'Session 8: Communicating Results'
---

# Publication Ready Figures
```{r}
# Loading required packages
library(tidyverse)

# Let's get our data
data_url <- "https://github.com/mickaeltemporao/CMT4A-CMSS-TEMPLATE/raw/main/data/ts_2016.rds"
raw_data <- readRDS(url(data_url,"rb"))

tmp_data <- raw_data %>%
  select(
    V161126,  # ideology
    V161010e,  # state 2 letters
    V163003,  # region
  ) %>%
  rename(
    ideology = V161126,
    state_code = V161010e,
    region = V163003,
  )

head(tmp_data)
```
## Filter and Recode values
```{r}

tmp_data <- tmp_data %>%
  filter(between(ideology, 1, 7)) %>%
  mutate(
    region = factor(region,  labels = c("Northeast", "Midwest", "South", "West")),
  )

head(tmp_data)

```

## Your Plot is the Message

```{r}
ggplot(tmp_data, aes(x=ideology)) +
  geom_bar()

```

# Advanced Figures

## Drawing Maps

To draw maps we will use the maps package.

```{r eval=FALSE}

# Install required packages
# Run line 9, if you get an error copy line 5 and run it in the console.
# install.packages('maps')

# Loading required packages
library(maps)

```


Once the *maps* package is loaded, we can easily draw maps!
```{r}
# Fetching the data
world_map <- map_data("world")

# Making a figure using geom_polygon
ggplot(world_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill="lightblue", colour = "white")
```


Let's make a map of a single country

```{r}

```



Let's look at the US

```{r}
# Create an object with the states map data
us_map <- map_data("state")
ggplot(us_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill="lightblue", colour = "white")

```


What does the map data look like?
```{r}
us_map

us_map %>% count(region)

```

The region variable seems to be the state. Let's rename that variable.

```{r}
us_map <- us_map %>% rename(state = region)

us_map
```

How does the data we used before look like?
```{r}
tmp_data
```

Let's create a new variable called "state_name" with the string names of states

```{r}
tmp_data <- tmp_data %>%
  mutate(
    state = case_when(
      state_code == "AK" ~ "Alaska",
      state_code == "AL" ~ "Alabama",
      state_code == "AR" ~ "Arkansas",
      state_code == "AZ" ~ "Arizona",
      state_code == "CA" ~ "California",
      state_code == "CO" ~ "Colorado",
      state_code == "CT" ~ "Connecticut",
      state_code == "DC" ~ "District of Columbia",
      state_code == "DE" ~ "Delaware",
      state_code == "FL" ~ "Florida",
      state_code == "GA" ~ "Georgia",
      state_code == "HI" ~ "Hawaii",
      state_code == "IA" ~ "Iowa",
      state_code == "ID" ~ "Idaho",
      state_code == "IL" ~ "Illinois",
      state_code == "IN" ~ "Indiana",
      state_code == "KS" ~ "Kansas",
      state_code == "KY" ~ "Kentucky",
      state_code == "LA" ~ "Louisiana",
      state_code == "MA" ~ "Massachusetts",
      state_code == "MD" ~ "Maryland",
      state_code == "ME" ~ "Maine",
      state_code == "MI" ~ "Michigan",
      state_code == "MN" ~ "Minnesota",
      state_code == "MO" ~ "Missouri",
      state_code == "MS" ~ "Mississippi",
      state_code == "MT" ~ "Montana",
      state_code == "NC" ~ "North Carolina",
      state_code == "ND" ~ "North Dakota",
      state_code == "NE" ~ "Nebraska",
      state_code == "NH" ~ "New Hampshire",
      state_code == "NJ" ~ "New Jersey",
      state_code == "NM" ~ "New Mexico",
      state_code == "NV" ~ "Nevada",
      state_code == "NY" ~ "New York",
      state_code == "OH" ~ "Ohio",
      state_code == "OK" ~ "Oklahoma",
      state_code == "OR" ~ "Oregon",
      state_code == "PA" ~ "Pennsylvania",
      state_code == "RI" ~ "Rhode Island",
      state_code == "SC" ~ "South Carolina",
      state_code == "SD" ~ "South Dakota",
      state_code == "TN" ~ "Tennessee",
      state_code == "TX" ~ "Texas",
      state_code == "UT" ~ "Utah",
      state_code == "VA" ~ "Virginia",
      state_code == "VT" ~ "Vermont",
      state_code == "WA" ~ "Washington",
      state_code == "WI" ~ "Wisconsin",
      state_code == "WV" ~ "West Virginia",
      state_code == "WY" ~ "Wyoming",
  )
)

tmp_data

```

The state in the us_map data is written with a lowercase. We need those columns to match!
```{r}
tmp_data <- tmp_data %>% mutate(state = tolower(state))
tmp_data
```

Let's get the average ideology by state
```{r}
state_idl <- tmp_data %>% group_by(state) %>% summarize(ideology = mean(ideology))
state_idl
```

We now have a dataset with a matching state column and the average ideology by state.
We can join these 2 datasets together and try to visualize the ideology by state.
```{r}
ideology_map <- us_map %>% left_join(state_idl)
```

# We can make a map!

```{r}
ggplot(ideology_map, aes(long, lat)) +
  geom_polygon(aes(fill = ideology, group = group), color = "white") +
  scale_fill_gradient2(
    low = "#377eb8", mid = "grey", midpoint = 4,
    high = "#e41a1c", limits = c(1, 7), breaks = c(1, 4 ,7)
  ) +
  theme_minimal()

```
I might not know all the state names let's add them!

```{r}
centroids <- us_map %>% group_by(state) %>% summarise(long = mean(long), lat = mean(lat))
centroids <- centroids %>% left_join(tmp_data)

 ggplot(ideology_map, aes(long, lat)) +
  geom_polygon(aes(fill = ideology, group = group), color = "white") +
  geom_text(data=centroids, aes(label = state_code, x = long, y = lat), size=3, color="white") +
  scale_fill_gradient2(
    low = "#377eb8", mid = "grey", midpoint = 4,
    high = "#e41a1c", limits = c(1, 7), breaks = c(1, 4 ,7)
  ) +
  theme_minimal()

```

Some like maps, some might find more classic figures easier to understand.

```{r}

ggplot(
  state_idl, aes(x = ideology, y = reorder(state, ideology))) +
  geom_vline(xintercept = 4, color = "black", alpha=.5) +
  geom_point(aes(color=ideology), size = 2) +
  geom_point(shape=1, size=2.5) +
  scale_x_continuous(
    limits = c(2,6),
    breaks = 2:6,
    labels = c("Liberal", "", "0\nModerate", "", "Conservative")
  ) +
  scale_color_gradient2(low="#377eb8", mid="grey", high="#e41a1c", midpoint=4) +
  labs(
    title = "Average Ideology by State",
    y = "State",
    x = "Liberal Conservative Self-placement",
    color = ""
  )

```


