---
title: "Session 7: Data Modeling (2/2)"
subtitle: "Learning Objectives:"
bibliography: references.bib
---

Demo 1: Improve plots - Themes and colors
Demo 2: Improve plots - Titles, subtitles, axis, caption

- colors palettes
- themes
- maps

- Learn to create and interpret regression tables
- Learn to make end to end research projects
- Retrospective

# Resources

## Readings

## Practice

```{r}

```


# Create a new variable called state_name with the string names of states
NESdta_short <- NESdta_short %>%
  mutate(state_name = case_when(state == 1~"Alabama", state == 2~"Alaska",
                                state == 4~"Arizona", state == 5~"Arkansas",
                                state == 6~"California", state == 8~"Colorado",
                                state == 9~"Connecticut", state == 10~"Delaware",
                                state == 11~"District of Columbia",
                                state == 12~"Florida", state == 13~"Georgia",
                                state == 15~"Hawaii", state == 16~"Idaho",
                                state == 17~"Illinois", state == 18~"Indiana",
                                state == 19~"Iowa", state == 20~"Kansas",
                                state == 21~"Kentucky", state == 22~"Louisiana",
                                state == 23~"Maine", state == 24~"Maryland",
                                state == 25~"Massachusetts", state == 26~"Michigan",
                                state == 27~"Minnesota", state == 28~"Mississippi",
                                state == 29~"Missouri", state == 30~"Montana",
                                state == 31~"Nebraska", state == 32~"Nevada",
                                state == 33~"New Hampshire", state == 34~"New Jersey",
                                state == 35~"New Mexico", state == 36~"New York",
                                state == 37~"North Carolina", state == 38~"North Dakota",
                                state == 39~"Ohio", state == 40~"Oklahoma",
                                state == 41~"Oregon",  state == 42~"Pennsylvania",
                                state == 44~"Rhode Island",  state == 45~"South Carolina",
                                state == 46~"South Dakota", state == 47~"Tennessee",
                                state == 48~"Texas", state == 49~"Utah",
                                state == 50~"Vermont", state == 51~"Virginia",
                                state == 53~"Washington", state == 54~"West Virginia",
                                state == 55~"Wisconsin", state == 56~"Wyoming"))

NESdta_short


# Hack Time!

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

1. Open a RMarkdown document and paste the following code in it: [my_project.Rmd](https://raw.githubusercontent.com/mickaeltemporao/CMT4A-CMSS/main/my_project.Rmd)
2. Open a Text file, save it as "references.bib" and paste the following code in it: [references.bib](https://raw.githubusercontent.com/mickaeltemporao/CMT4A-CMSS/main/references.bib)


#### References

